---
title: "coursera practical machine learning” course project"
author: "Yue Guan"
date: "22/05/2020"
output: html_document
---

## Exploratory Data Analysis

### Data import and util packages installation


```{r message=FALSE, warning=FALSE}


library(dplyr)
library(magrittr)
library(caret)
library(doParallel)
library(tidyr)
library(corrplot)
library(randomForest)
library(RColorBrewer)
library(e1071)


train_data <- read.csv("./pml-training.csv", stringsAsFactors = FALSE)
validation_data <- read.csv("./pml-testing.csv", stringsAsFactors = FALSE)
```

### Dataset description

```{r message=FALSE, warning=FALSE}
#Convert the data to a data frame table type to simplify visualization
tbl_trainset<-tbl_df(train_data)
class(tbl_trainset)
tbl_trainset
#Find out columns with 0 data inside for cleaning
tbl_trainset %>%
  select(everything()) %>%  
  summarise_all(funs(sum(is.na(.)|is.null(.))))


```



### Train test repartition

Here I use 70% of train set as training data

```{r message=FALSE, warning=FALSE}
inTrain <- createDataPartition(train_data$classe, p=0.7, list=FALSE)
trainSet <- train_data[inTrain, ]
testSet <- train_data[-inTrain, ]
dim(trainSet)
dim(testSet)
dim(trainSet)
```

### Data clean

Remove label columns, zero variables and variables which 90% or more data are N/A

```{r message=FALSE, warning=FALSE}
#Remove name and time stamp variables
trainSet <- trainSet[, -(1:5)]
testSet <- testSet[, -(1:5)]

#find zero variables
zeros <- nearZeroVar(trainSet)
# remove
trainSet <- trainSet[,-zeros]
testSet <- testSet[,-zeros]

#find na variables
nas <- sapply(trainSet, function(val) mean(is.na(val))) > 0.9
#remove 
trainSet <- trainSet[, nas==FALSE]
testSet <- testSet[, nas==FALSE]
dim(trainSet)
```
## Visualize correlation matrix


```{r warning=FALSE, error=FALSE}
corrMatrix <- cor(trainSet[, -54])
corrplot(corrMatrix, method = "color", tl.cex = 0.5)
```

# Building Prediction Model  with SVM, knn and random forest

## Building Prediction Model  with SVM
```{r}
svmCtrl <- trainControl(method="cv", number = 3, verboseIter = FALSE)
svmModel <- train(classe ~ ., data=trainSet, method="svmLinear", trControl=svmCtrl)
svmModel$finalModel
```
Now perform prediction on test dataset and then create confusion matrix
```{r warning=FALSE, error=FALSE}
# do prediction on testSet
svmPrediction <- predict(svmModel, newdata = testSet)
# create confusion matrix
svmConfMat <- confusionMatrix(table(svmPrediction, testSet$classe))
svmConfMat
```
Plot confusion matrix for prediction result
```{r warning=FALSE, error=FALSE}
plot( svmConfMat$table, col = svmConfMat$byClass, main=paste("Prediction Accuracy with svm: ", round(svmConfMat$overall['Accuracy'], 4) ))
```


## Building Prediction Model  with K-nearest neighbor
```{r}
knnCtrl <- trainControl(method="cv", number = 3, verboseIter = FALSE)
knnModel <- train(classe ~ ., data=trainSet, method="knn", trControl=knnCtrl)
knnModel$finalModel
```
Now perform prediction on test dataset and then create confusion matrix
```{r warning=FALSE, error=FALSE}
# do prediction on testSet
knnPrediction <- predict(knnModel, newdata = testSet)
# create confusion matrix
knnConfMat <- confusionMatrix(table(knnPrediction, testSet$classe))
knnConfMat
```
Plot confusion matrix for prediction result
```{r warning=FALSE, error=FALSE}
plot( knnConfMat$table, col = knnConfMat$byClass, main=paste("Prediction Accuracy with knn: ", round(knnConfMat$overall['Accuracy'], 4) ))
```

## Building Prediction Model  with random forest

```{r warning=FALSE, error=FALSE}
rfCtrl <- trainControl(method="cv", number = 3, verboseIter = FALSE)
rfModel <- train(classe ~ ., data=trainSet, method="rf", trControl=rfCtrl)
rfModel$finalModel
```

Now perform prediction on test dataset and then create confusion matrix
```{r warning=FALSE, error=FALSE}
# do prediction on testSet
rfPrediction <- predict(rfModel, newdata = testSet)
# create confusion matrix
rfConfMat <- confusionMatrix(table(rfPrediction, testSet$classe))
rfConfMat
```

Plot confusion matrix for prediction result
```{r warning=FALSE, error=FALSE}
plot( rfConfMat$table, col = rfConfMat$byClass, main=paste("Prediction Accuracy with RandomForest: ", round(rfConfMat$overall['Accuracy'], 4) ))
```
```

We got a better accuracy with RandomForest model

# Check RF model on validation dataset for quiz
```{r warning=FALSE, error=FALSE}
predictions <- predict(rfModel, newdata = validation_data)
predictions

```


